\chapter{Basics}

\section{Definitions}

\begin{definition}[\cindex{topology}]
    A topology on a set $X$ is a collection $\mathcal{T}$ of subsets of $X$ having the following property:
    \begin{enumerate}
        \item $\emptyset$ and $X$ are in $\mathcal{T}$.
        \item The union of any subcollection of $\mathcal{T}$ is in $\mathcal{T}$.
        \item the intersection of finite subcollection of $\mathcal{T}$ is in $\mathcal{T}$.
    \end{enumerate}

The set $X$ with a topology is called \cindex{topological space}. The set $U$ in $\mathcal{T}$ is called \cindex{open set}.
\end{definition}

\begin{definition}[\cindex{closed set}]
    A subset $A$ of a topological space $X$ is closed if the set $X - A$ is open.
\end{definition}

\begin{definition}
    Let $X$ be a topological space. Then:
    \begin{enumerate}
        \item $\emptyset$ and $X$ are closed
        \item Arbitrary union of open sets are open
        \item Arbitrary intersections of closed sets are closed
        \item Finite unions of closed sets are closed
        \item Finite intersection of open sets are open
    \end{enumerate}
\end{definition}


\begin{definition}[\cindex{interior}]
    Given a subset $A$ of a topological space $X$, the interior $\interior{A}$ of $A$ is defined as the union of all open sets contained in $A$.
\end{definition}

\begin{definition}[\cindex{closure}]
    the closure $\closure{A}$ of $A$ is defined as the intersection of all closed sets containing $A$. In subspace topology it maybe ambiguous to express for which the closure it. We will use $\closure{A}_X$ to mean it is a closure in topology $X$.
\end{definition}

\begin{theorem}
    The relationship of interior and closure is
\begin{equation}
    \mathring{A} \subset A \subset \closure{A}
\end{equation}
\end{theorem}



\begin{definition}[\cindex{boundary}]
    The boundary of $A$, denoted by $\boundary{A}$, is defined as
    \begin{equation}
        \boundary{A} = \closure{A} \cap \closure{X - A}
    \end{equation}
\end{definition}


\begin{definition}[\cindex{neighborhood}]
    A neighborhood of $x$ is an open set $U$ containing $x$.
\end{definition}


\begin{definition}[\cindex{limit point}]
    Let $A$ be a subset of topological space $X$. $x$ is a limit point of $A$ if every neighborhood of $x$ intersects $A$ in some point \emph{other than $x$}.
\end{definition}



\begin{theorem}
    Here are the list of properties around interior, closure and boundary:
    \begin{itemize}
        \item $x \in \overline{A}$ if and only if every open set $U$ containing $x$ intersects $A$
        \item $\interior{A}$ is the largest open set contained in $A$
        \item $\closure{A}$ is the smallest closed set that contains $A$
        \item $Y \cap \closure{A}_X = \closure{A}_Y$
        \item $Y \cap \interior{A}_X \subseteq \interior{A}_Y$
        \item $x$ is a limit point of $A$ if and only if every open set $U$ containing $x$ intersects $A$ other than $x$
        \item $x$ is a limit point of $A$ if it belongs to $\closure{A - \set{x}}$
        \item A limit point $x$ of $A$ may or may not in $A$
    \end{itemize}
\end{theorem}
\begin{proof}
    For the counterexample of $Y \cap \interior{A}_X \subseteq \interior{A}_Y$, let $A = [0,1)$, $Y = [0,2]$. $\interior{A}_X = (0,1)$ while $\interior{A}_Y = [0,1)$.
\end{proof}





\begin{theorem}
    Let $A'$ be the set of all limit points of $A$. Then $\closure{A} = A \cup A'$.
\end{theorem}
\begin{proof}
    Let $x \in \closure{A}$. $\forall U, x \in U, U \cap A = C \neq \emptyset$, so either $x \in C \Rightarrow x \in A$, or $x \notin C \Rightarrow x \in A'$.
\end{proof}

\begin{theorem}
A subset of $A$ is closed if and only if it contains all its limit points. So if a subset does not have a limit point, it is closed.
\end{theorem}


\begin{definition}[\cindex{dense}]
    A subset $E$ of a topological space is dense if $\closure{E} = X$
\end{definition}

\begin{definition}[\cindex{separable}]\label{separable}
    A topological space is separable if it has a countable dense subset.
    
    Note: separable is different from separation (see \defiref{separation})
\end{definition}


\begin{definition}[\cindex{order topology}]
    Let $X$ be a set with a simple order relation. Let $\mathcal{B}$ be the collection of all sets:
    \begin{enumerate}
        \item All open intervals $(a,b)$ in $X$.
        \item All intervals $[a_0, b)$ where $a_0$ is the smallest element (if any) of $X$.
        \item All intervals $(a, b_0]$ where $b_0$ is the largest element (if any) of $X$.
    \end{enumerate}
    The topology generated by basis $\mathcal{B}$ is called the order topology.
\end{definition}


% basis
\section{Basis}

\begin{definition}[\cindex{basis}]
    A basis for a topology is a collection $\mathcal{B}$ of subsets of $X$ that:
    \begin{enumerate}
        \item For each $x \in X$, there is at least one basis element $B$ containing $x$.
        \item For $B_1, B_2 \in \mathcal{B}$, if $x \in B_1 \cap B_2$, there is a basis element $B_3$ that $x \in B_3 \subset B_1 \cap B_2$.
    \end{enumerate}
\end{definition}

The topology $\mathcal{T}$ generated by basis $\mathcal{B}$ is: a subset $U$ of $X$ is open in $X$ if for each $x \in U$, there is a basis element $B \in \mathcal{B}$ such that $x \in B \subset U$.


\begin{definition}[\cindex{subbasis}]
    A subbasis $\mathcal{S}$ for a topology on $X$ is a collection of subsets of $X$ whose union equals $X$. The topology generated by the subbasis $\mathcal{S}$ is the collection $\mathcal{T}$ of all unions of finite intersections of elements of $\mathcal{S}$.
\end{definition}


\begin{example}
    Common topologies:
\begin{itemize}
    \item \cindex{discrete topology}: the collection of all subsets of $X$.
    \item \cindex{finite complement topology}: a collection of subsets $U$ of $X$ that $X - U$ is either finite or all of $X$.
    \item \cindex{standard topology} on order set: a collection of all open intervals $(a,b) = \set{x | a < x < b}$.
    \item \cindex{lower limit topology} \cindex{$\mathcal{R}_l$}: $[a,b) = \set{x | a \leq x < b}$.
    \item \cindex{K-topology}: let $K = \set{\displaystyle \frac{1}{n} | n \in \mathbb{Z}_{+}}$. The basis are the collection of all open intervals $(a,b)$ and $(a,b) - K$.
\end{itemize}
\end{example}





\begin{theorem}
    Let $\mathcal{B}$ be a basis for a topology $\mathcal{T}$ of $X$. Then $\mathcal{T}$ equals the collection of all unions of elements of $\mathcal{B}$.
\end{theorem}

\begin{theorem}
    Suppose $\mathcal{C}$ is a collection of open sets of $X$ such that for each open set $U$ of $X$ and $x \in U$, there is an element $C$ of $\mathcal{C}$ that $x \in C \subset U$. Then $\mathcal{C}$ is a basis for the topology of $X$.    
\end{theorem}


How to understand the relation among topology, basis and subbasis:
\begin{itemize}
    \item topology is closed with arbitrary union and finite intersection
    \item basis only needs to do arbitrary union
    \item subbasis has to do finite intersection
\end{itemize}



% subspace topology
\section{Subspace Topology}

\begin{definition}[\cindex{subspace topology}]
    Let $(X, \mathcal{T})$ be a topology. If $Y$ is a subset of $X$, the collection $\mathcal{T}_{Y} = \set{Y \cap U | U \in \mathcal{T}}$ is a topology on $Y$ which is called the subspace topology.
\end{definition}

\begin{theorem}
    If $\mathcal{B}$ is a basis for the topology of $X$ then the collection $\mathcal{B}_Y = \set{B \cap Y | B \in \mathcal{B}}$ is a basis for the subspace topology on $Y$.
\end{theorem}

\begin{theorem}
    Let $Y$ be a subspace of $X$. If $U$ is open in $Y$ and $Y$ is open in $X$, then $U$ is open in $X$.    
\end{theorem}

The subspace topology $Y$ need not be the same as the topology in $X$. For example, for $Y=[0,1]$ as the subspace of order topology on $\realnumber$. $[0,a)$ is not open on $\realnumber$ but open on $Y$.


\begin{theorem}[\cindex{convex}]
    A subset $Y$ of $X$ is convex in $X$ if for each pair of points $a < b$ of $Y$, the entire interval $(a,b)$ of points of $X$ is in $Y$.     
\end{theorem}

\begin{theorem}
    Let $X$ be an ordered set in order topology, and let $Y$ be a subset of $X$ that is convex in $X$. Then the order topology on $Y$ is the same as the topology inherits as a subspace of $X$.    
\end{theorem}


\begin{example}[\cindex{ordered square}]
    Let $I=[0,1]$. The dictionary order on $I \times I$ is called ordered square, and denoted as \cindex{$I_o^2$}. 
\end{example}





% hausdorff
\section{Hausdorff Space}

\begin{definition}[\cindex{Hausdorff space}]
    A topological space $X$ is Hausdorff space if for each pair $x_1, x_2$ of distinct points of $X$, there exists neighborhood $U_1$ and $U_2$ of $x_1$ and $x_2$ that are disjoint.
\end{definition}

\begin{definition}[\cindex{$T_1$} axiom]
Every finite point set is closed.    
\end{definition}

\begin{theorem}
    Hausdorff space is $T_1$.
\end{theorem}
\begin{proof}
    For a single point $x$, check what is its closure. for any $y \neq x$, there is $y \in U$ and $x \in V$ that $U \cap V = \ \emptyset$, so $y$ has a neighbor that did not contain $x$, so $y$ is not in the closure of $x$. Since no other point is in the closure of $x$, $x$ is the closure of itself.
\end{proof}


\begin{theorem}
Let $X$ be a space satisfying the $T_1$ axiom. Let $A$ be a subset of $X$. Then the point $x$ is a limit point of $A$ if and only if every neighborhood of $x$ contains infinite many points of $A$.    
\end{theorem}
\begin{proof}
    Assume the neighborhood $U$ of $x$ intersects on finite points, which are $\set{x_i}$. The set $X - \cup_n \set{x_i}$ is open, so $U \cap (X - \cup_n \set{x_i})$ is open and did not intersect $A$.
\end{proof}


\begin{definition}[\cindex{converge}]
    A sequence $x_1, x_2, ... $ of points of space $X$ converges to point $x$ of $X$ if for each neighborhood $U$ of $x$, there is a positive integer $N$ such that $x_n \in U$ for all $n \geq N$.
\end{definition}

\begin{theorem}
If $X$ is a Hausdorff space, then a sequence of points of $X$ converges to at most one point of $X$.    
\end{theorem}

\begin{theorem}
Properties of Hausdorff space:
\begin{itemize}
    \item Every simple ordered set is Hausdorff space in order topology.
    \item The produce of two Hausdorff space is a Hausdorff space.
    \item A subspace of a Hausdorff space is a Hausdorff space. 
\end{itemize}
\end{theorem}



\section{Continuous Functions}


\begin{definition}[\cindex{continuous}]
    A function $f: X \rightarrow Y$ is continuous if for each open subset $V$ of $Y$, the set $f^{-1}(V)$ is an open subset of $X$.
    
    So continuity depends not only on $f$, but also on the topology of domain and range spaces.
\end{definition}

To prove a function $f$ is continuous, it is suffice to prove the inverse image of subbasis is open.

\begin{theorem}
Let $f: X \rightarrow Y$. Then the following are equivalent:
\begin{enumerate}
    \item $f$ is continuous.
    \item For every subset $A$ of $X$, one has $f(\closure{A}) \subset \closure{f(A)}$.
    \item For every closed set $B$ of $X$, the set $f^{-1}(B)$ is closed in $X$.
    \item For each $x\in X$ and each neighborhood $V$ of $f(x)$, there is a neighborhood $U$ of $x$ such that $f(U) \subset V$.
\end{enumerate}    

If the condition (4) holds for a point $x$ of $X$, then $f$ is continuous at the point $x$.
\end{theorem}
\begin{proof}
    For (1) $\rightarrow$ (2), $\forall y \in f(\closure{A}), \exists x \in \closure{A}, y = f(x)$. Take $U$ that $y \in U$, there is $V = f^{-1}(U)$ that is open and $x \in V$. Since $x \in \closure{A}$, $\exists t \in V \cap A$. So $f(t) \in U \cap f(A) \neq \emptyset$.
    
    For (2) $\rightarrow$ (3), $f(\closure{A}) \subset \closure{f(A)} = B$, so $f^{-1} \left(f(\closure{A}) \right) \subset f^{-1}(B) = A$. So $\closure{A} \subset f^{-1} \left(f(\closure{A}) \right) = A$.
\end{proof}


\begin{definition}[\cindex{homeomorphism}]
    Let $f:X \rightarrow Y$ be a bijection. if $f$ and $f^{-1}$ are continuous, $f$ is called homeomorphism and the map $f$ is a \cindex{topological imbedding}, or \cindex{imbedding} of $X$ in $Y$.
\end{definition}


\begin{theorem}[\cindex{pasting lemma}]
    Let $X = A \cup B$ where $A$ and $B$ are closed in $X$. Let $f: A \rightarrow Y$ and $g: B \rightarrow Y$ be continuous. If $\forall x \in A \cap B, f(x) = g(x)$, then $f$ and $g$ combine to give a continuous function $h: X \rightarrow Y$, defined by $\forall x \in A, h(x) = f(x) $ and $\forall x \in B , h(x) = g(x)$.
\end{theorem}

\begin{theorem}[\cindex{coordinate functions}]
Let $f: A \rightarrow X \times Y$ defined as $f(a) = \left(f_1(a), f_2(a) \right)$. Then $f$ is continuous if and only if $f_1$ and $f_2$ are continuous. The map $f_1$ and $f_2$ are called the coordinate functions of $f$.
\end{theorem}


% product space
\section{Product Space}

\begin{definition}[\cindex{cartesian product}]
    Let $\set{A_{\alpha}}_{\alpha \in \mathcal{J}}$ be an indexed family of sets. The cartesian product $\displaystyle \prod_{\alpha \in \mathcal{J}} A_{\alpha}$ of this indexed family is defined to be the set of all functions $\set{f : \mathcal{J} \rightarrow \displaystyle \bigcup_{\alpha \in \mathcal{J}} A_{\alpha}}$ that $f(\alpha) \in A_{\alpha}$.
\end{definition}

\begin{definition}[\cindex{box topology}]
    Let $\set{X_{\alpha}}_{\alpha \in \mathcal{J}}$ be an index family of topological spaces. The box topology generated by the basis of all sets of the form $\displaystyle \prod_{\alpha \in \mathcal{J}} U_{\alpha}$ where $U_{\alpha}$ is open (or a basis) in $X_{\alpha}$. 
\end{definition}

\begin{definition}[\cindex{projection mapping}]
    A projection mapping $\pi_\beta$ is defined as $\pi_\beta \left( (x_{\alpha})_{\alpha \in \mathcal{J}} \right) = x_\beta$.
\end{definition}

\begin{definition}[\cindex{product topology}]
    Define $\mathcal{S}_{\beta} = \set{\pi_{\beta}^{-1}(U_{\beta}) | U_{\beta}\text{ is open in }X_{\beta}}$ and let $\mathcal{S} = \displaystyle \bigcup_{\beta \in \mathcal{J}} \mathcal{S}_{\beta}$. The topology generated by the subbasis $\mathcal{S}$ is called the product topology. 
    
    So the product topology on $\displaystyle \prod_{\alpha \in \mathcal{J}} X_{\alpha}$ has the basis all sets of the form $\displaystyle \prod_{\alpha \in \mathcal{J}} U_{\alpha}$, where $U_{\alpha}$ is open in $X_{\alpha}$ and $U_{\alpha} = X_{\alpha}$ except for finite many value of $\alpha$.
\end{definition}

\begin{theorem}
Let $A_{\alpha} \subset X_{\alpha}$. $\displaystyle \prod \closure{A_{\alpha}} = \closure{\prod A_{\alpha}}$ in either box topology or product topology.
\end{theorem}


\begin{theorem}
Let $\displaystyle f : A \rightarrow \prod_{\alpha} X_{\alpha}$ be given by $f(a) = \left( f_{\alpha}(a) \right)$ where $f_{\alpha} : A \rightarrow X_{\alpha}$. In product topology $f$ is continuous if and only if $f_{\alpha}$ is continuous.

This is the reason why product topology is preferred over box topology.
\end{theorem}

\begin{example}[counterexample in box topology where $f$ is not continuous]
    Define $f: \realnumber \rightarrow \realnumber^{\omega}$ by $f(t) = (t,t,t, \cdots)$. Define a open set in box topology $B = (-1, 1) \times (- \frac{1}{2}, \frac{1}{2}) \times  (- \frac{1}{3}, \frac{1}{3}) \times \cdots$. $f^{-1}(B)$ is not open in $R$.
\end{example}

\section{Metric Topology}

\begin{definition}[\cindex{metric}]
    A metric on a set $X$ is a function $d: X \times X \rightarrow \realnumber$ with the following property:
    \begin{enumerate}
        \item $d(x,y) \geq 0$.
        \item $d(x,y) = 0 \Leftrightarrow x = y$.
        \item $d(x,y) = d(y,x)$.
        \item $d(x,y) + d(y,z) \geq d(x,z)$.
    \end{enumerate}
    
    $d(x,y)$ is also called the \cindex{distance} between $x$ and $y$.
\end{definition}

\begin{definition}[\cindex{$\epsilon$-ball}]
    For a metric space $X$ with distance function $d$, the $\epsilon$-ball centered at $x$ is defined as $B_d(x, \epsilon) = \set{y | d(x,y) < \epsilon}$.
\end{definition}

\begin{definition}[\cindex{metric topology}]
    For a metric space $X$ with distance function $d$, the metric topology is defined by taking as basis all the $\epsilon$-balls $B_d(x,\epsilon)$ for all $x$ and $\epsilon > 0$. A topological space $X$ is \cindex{metrizable} if there exists a metric $d$ on $X$ that induces the same topology.
\end{definition}

\begin{definition}[\cindex{bounded}]
    A subset $A$ of $X$ is bounded if $\exists m, \forall a_1, a_2 \in A: d(a_1, a_2) \leq m$. If $A$ is bounded and nonempty, the \cindex{diameter} of $A$ is defined as $\text{diam } A = \supremum{d(a_1, a_2)| a_1, a_2 \in A}$.
    
    Bounded is not a topology property because it depends on the value of $d$.
\end{definition}

\begin{definition}[\cindex{standard bounded metric}]
    Let $X$ be a metric space with metric $d$. The standard bounded metric corresponding to $d$ is defined as $\closure{d}(x,y) = \text{min}\set{d(x,y), 1} $.
\end{definition}

\begin{definition}[\cindex{norm}]
    For $x \in \realnumber^n$, the norm of $x$ is $\norm{x} = \sqrt{x_1^2 + ... + x_n^2}$. The \cindex{euclidean metric} $d$ on $\realnumber^n$ is defined as $d(x,y) = \norm{x - y}$. The \cindex{square metric} $\rho$ is defined as $\rho (x,y) = \text{max}\set{\absolutevalue{x_1 - y_1}, ... , \absolutevalue{x_n - y_n} }$.
\end{definition}

\begin{definition}[\cindex{uniform metric}]
    When $n \rightarrow \infty$, the euclidean and square metric fail to work. So need to define a metric that works for any $\mathcal{J}$. Define $\closure{d}(x,y) = \text{min}\set{\absolutevalue{x - y} , 1}$ and the uniform metric $\closure{\rho}(x,y) = \supremum{ \closure{d}(x_{\alpha}, y_{\alpha}) | \alpha \in \mathcal{J}} $. The topology generated by uniform metric is called \cindex{uniform topology}.
\end{definition}
\begin{theorem}
Let $f: X \rightarrow Y$. Let $X$ and $Y$ be metrizable with $d_X$ and $d_Y$. $f$ is continuous is equivalent to the requirement that $\forall x \in X, \forall \epsilon > 0, \exists \delta > 0, d_X (x,y) < \delta \Rightarrow d_Y \left(f(x), f(y) \right) < \epsilon $.
\end{theorem}


\begin{definition}[\cindex{first countability axiom}]\label{first_countability_axiom}
    A space $X$ that has a countable basis at each of its points is said to satisfy the first countability axiom. 
    
    A metric space always satisfies the first countability axiom.
\end{definition}

\begin{theorem}[sequence lemma]
    Let $A \subset X$. If there is a sequence of points of $A$ converging to $x$, then $x \in \closure{A}$. The converse holds if $X$ is metrizable.
    
    Note: the sequence lemma is quite useful in proving a topology space is not metrizable.
\end{theorem}
\begin{proof}
    Use \defiref{first_countability_axiom}. For a $x$, choose neighbor $B_d(x, \frac{1}{n})$, and let it intersect with $A$.
\end{proof}


\begin{theorem}
    Let $f: X \rightarrow Y$. If $f$ is continuous, then for every convergent sequence $x_n \rightarrow x$ in $X$, the sequence $f(x_n) \rightarrow f(x)$. The converse holds if $X$ is metrizable ($Y$ does not need to be metrizable).
\end{theorem}
\begin{proof}
    For a $A \subset X$,  $x_n \rightarrow x \Rightarrow x \in \closure{A}$. $f(x_n) \rightarrow f(x) \Rightarrow f(x) \in \closure{f(A)}$. What we need is a way to find a sequence of $x_n$, and use \defiref{first_countability_axiom}.
\end{proof}




\begin{definition}[\cindex{converge uniformly}]
    A sequence $f_n$ converge uniformly to $f$ if $\forall \epsilon > 0,\forall x \in X, \exists N \in \mathbb{Z}_{+}, \forall n > N, d \left( f_n (x) , f(x) \right) < \epsilon$.
\end{definition}

\begin{theorem}
    Let $f_n: X \rightarrow Y$ be a sequence of continuous functions from $X$ into metric space $Y$. If $f_n$ converges uniformly to $f$, then $f$ is continuous. (no requirement on $X$)
\end{theorem}
\begin{proof}
    First find a $N$ that $f_n$ and $f$ are close enough. Then for $f_n$, find a $\epsilon$-ball small enough to restrict how big the $f$ is.
\end{proof}





% metrizable of R^j

\section{Metrizable of $\realnumber^J$}


\begin{theorem}
The uniform topology on $\realnumber_{\mathcal{J}}$ is finer than the product topology and coarser than the box topology. They are all different if $\mathcal{J}$ is infinite.
\end{theorem}
\begin{proof}
    For a basis in product topology, there are finite $B_i$ that is not $R$. Find the minimum $\epsilon$ and make $\closure{\rho} < \epsilon$. For the box topology, find basis that is half of the size of $\closure{\rho}$.
\end{proof}



\begin{theorem}[$\realnumber^{\omega}$ is metrizable]
    Define $\displaystyle D(x,y) = \supremum{\frac{\closure{d}(x_i, y_i )}{i}} $. Then $D$ is a metric that induces the product topology on $\realnumber^\omega$.
\end{theorem}
\begin{proof}
    (the metric is changed from $\closure{\rho}(x,y) = \supremum{ \closure{d}(x_{\alpha}, y_{\alpha})} $ to $\supremum{\frac{\closure{d}(x_i, y_i )}{i}} $) First need to prove $D$ is a metric by proving the triangle inequality. The trick is to choose a $N$ that $\frac{1}{N} < \epsilon$ so all $\realnumber$ is included in the $\epsilon$-ball.
\end{proof}


\begin{example}[$\realnumber^\omega$ is not metrizable in box topology]
    Let $A = \set{(x_1, x_2, ... ) | x_i > 0}$. We have $(0,0,0,...) \in \closure{A}$. For a sequence $a_n = (x_{n_1}, x_{n_2}, ...)$, the open set $\set{(- x_{1_1}, x_{1_1}) \times (- x_{2_2}, x_{2_2}), ...}$ contains no $a_i$.
\end{example}


\begin{example}[$\realnumber^{\mathcal{J}}$ is not metrizable in product topology]
    Assume $\mathcal{J}$ is uncountable. Let $A$ contains all $x$ that $\pi_{\beta}(x) = 1$ for all but finite many $\beta$. Then $(0,0,...)$ is in $\closure{A}$.
    
    Assume there is a sequence $a_n$ that converge to $(0,0,...)$. Let $J_n = \set{\beta | \pi_\beta (a_n) \neq 1 }$. Then $\Sigma_n \absolutevalue{J_n}$ is countable of finite value, which is countable. Because $\mathcal{J}$ is uncountable, there exists a $\gamma$ that $\forall i, \pi_\gamma (a_i) = 1$. Take a basis $\pi_{\gamma}^{-1}(-1,1)$, which contains $(0,0,...)$ but intersect with no $a_i$.
\end{example}










